---
documentclass: jss
author:
  - name: David U. Garibay-Treviño
    affiliation: |
      | Center for Research and 
      | Teaching in Economics (CIDE)
    affiliation2: Center for Research and Teaching in Economics 
    address: |
      | First line
      | Second line
    email: \email{david.garibay@cide.edu}
    url: http://rstudio.com
  - name: Fernando Alarid-Escudero
    affiliation: |
      | Center for Research and 
      | Teaching in Economics (CIDE)
title:
  formatted: "\\pkg{demogmx}: An R mexican demographic information package"
  # If you use tex in the formatted title, also supply version without
  plain:     "A mexican demographic information Package demogmx"
  # For running headers, if needed
  short:     "\\pkg{demogmx}: A Mexican demographic Package"
abstract: >
  With information from CONAPO and INEGI  we made a new package that brings data 
  about births, migration, population, deaths and aging rate for each mexican 
  state as well as for the entire country. 
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
library(rmarkdown)
```

# Introduction

This template demonstrates some of the basic LaTeX that you need to know to create a JSS article.

## Code formatting

In general, don't use Markdown, but use the more precise LaTeX commands instead:

* \proglang{Java}
* \pkg{plyr}

One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax).

If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute. You can also provide a custom identifier if necessary. See the header of Section \ref{r-code} for example.

# \proglang{R} code {short-title="R code" #r-code}

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```

## Features specific to \pkg{rticles} {short-title="Features specific to rticles"}

* Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409).
* Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format.

\begin{small}
\begin{equation}\label{eq:si_hp_ode}
\begin{split}
\frac{dP_{1, j}}{dt} & = b_{j}(t) - \left(d_{1, j}(t)+ \eta_{1, j}(t) + \theta_{1, j}(t) + \mu_{1, j}(t) + \mu_{1, j}^{H}(t)\right) P_{1, j},  \hspace{0.2cm} j = male, female\\
\frac{dP_{i, j}}{dt} & = d_{i-1, j}(t)P_{i-1, j} - \left(d_{i, j}(t) + \eta_{i, j}(t) +         \theta_{i, j}(t) + \mu_{i, j}(t) + \mu_{i, j}^{H}(t)\right) P_{i, j}, \hspace{0.2cm} i=2,\ldots,n \hspace{0.1cm} ; \hspace{0.1cm} j = male, female \\
\frac{dDOC_{i, j}}{dt} & = \mu_{i, j}(t) P_{i, j}, \hspace{0.2cm} i = 1,\ldots, n \hspace{0.2cm} and \hspace{0.2cm} j = male, female \\ %\text{ for all } i = 1, \ldots, n,\\
\frac{dDH_{i, j}}{dt} & = \mu^{H}_{i, j}(t) P_{i, j}, \hspace{0.2cm} i = 1, \ldots, n  \hspace{0.2cm} and \hspace{0.2cm} j = male, female %\text{ for all } i = 1, \ldots, n,\\
\end{split}
\end{equation}
\end{small}

\begin{small}
\begin{equation}\label{eq:si_hp_ode}
\begin{split}
\frac{dP_{1, j}}{dt} & = b_{j}(t) - \left(d_{1, j}(t)+ \eta_{1, j}(t) + \theta_{1, j}(t) + \mu_{1, j}(t) + \mu_{1, j}^{H}(t)\right) P_{1, j} \\
\frac{dP_{i, j}}{dt} & = d_{i-1, j}(t)P_{i-1, j} - \left(d_{i, j}(t) + \eta_{i, j}(t) +         \theta_{i, j}(t) + \mu_{i, j}(t) + \mu_{i, j}^{H}(t)\right) P_{1, j} \\
\frac{dDOC_{i, j}}{dt} & = \mu_{i, j}(t) P_{1, j} \\
\frac{dDH_{i, j}}{dt} & = \mu^{H}_{i, j}(t) P_{1, j} \\
\end{split}
\end{equation}
\end{small}

where $P_{i, j}$ is the population in age group $i$ and sex group $j$, where $DOC_{i, j}$ is the number of deaths from other causes in age group $i$ and sex group $j$, $DH_{i, j}$ is the number of homicides in age group $i$ and sex group $j$, $d_{i, j}(t)$ is the rate of aging from age group $i$ to age group $i+1$, $\mu_{i, j}(t)$ is the background mortality for age group $i$ and sex group $j$ in year $t$ and $\mu_{i, j}^{H}(t)$ is the homicide rate for age group $i$ and sex group $j$ in year $t$, $\eta_{i, j}$ represents migration rate for age group $i$ and sex group $j$, while $\theta_{i, j}$ is the immigration rate for age group $i$ and sex group $j$.



$$\frac{dP_{1, j}}{dt} = births_j + (immigration - emmigration - aging_{i, j} -  background\ mort)P_{1,j}$$

$$\frac{dP_{i, j}}{dt} = (aging_{i-1, j})P_{(i-1), j} + (immigration - emmigration - aging_{i,j} - background\ mort)P_{i,j}$$

